---
title: "DMD_vs_Healthy_Muscle"
author: "Kobe Ikegami"
date: "2023-10-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1
```{r}
library(SeuratObject)
library(Seurat)
library(dplyr)
library(Seurat)
library(patchwork)
library(tidyverse)
```

```{r warning=FALSE}
#Import data and create objects: for each sample
dmd.data <- Read10X(data.dir = "/Users/ikegamik/Documents/kikegami/side_projects/tlr7_hunt/dmd")
cntl.data <- Read10X(data.dir = "/Users/ikegamik/Documents/kikegami/side_projects/tlr7_hunt/cntl")

dmd <- CreateSeuratObject(counts = dmd.data, project = "dmd", min.cells = 3, min.features = 200)

cntl <- CreateSeuratObject(counts = cntl.data, project = "cntl", min.cells = 3, min.features = 200)


```

Searches for mitochondrial RNA based on the "Mt" tag (human = "MT" or mouse = "mt")
```{r}
dmd[["percent.mt"]] <- PercentageFeatureSet(dmd, pattern = "^MT-")
cntl[["percent.mt"]] <- PercentageFeatureSet(cntl, pattern = "^MT-")
```


```{r}
#plot counts for rna, features, and mitochondrial RNA to begin filtering out outlier cells
plot1N <- FeatureScatter(dmd, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2N <- FeatureScatter(cntl, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot1N + plot2N
plot1N <- FeatureScatter(dmd, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot2N <- FeatureScatter(cntl, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1N + plot2N

VlnPlot(dmd, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
VlnPlot(cntl, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

```

```{r}
#APPLY FILTERS
dmd_sub <- subset(dmd, subset = nFeature_RNA < 2000 & nFeature_RNA > 200 & nCount_RNA < 4000 & percent.mt < 16)
cntl_sub <- subset(cntl, subset = nFeature_RNA < 1500 & nFeature_RNA > 200 & nCount_RNA < 3200 & percent.mt < 16)

```

```{r}
#Replot filtered data and see if stricter thresholds are necessary
plot1N <- FeatureScatter(dmd_sub, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2N <- FeatureScatter(cntl_sub, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot1N + plot2N
plot1N <- FeatureScatter(dmd_sub, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot2N <- FeatureScatter(cntl_sub, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1N + plot2N

VlnPlot(dmd_sub, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
VlnPlot(cntl_sub, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

```
normalize data and model mean variance in cell expression. Then do differential expression within treatments to find DE features
```{r warning=FALSE}
dmd_sub <- NormalizeData(dmd_sub)
cntl_sub <- NormalizeData(cntl_sub)

dmd_sub <- FindVariableFeatures(dmd_sub, selection.method = "vst", nfeatures = 2000)
cntl_sub <- FindVariableFeatures(cntl_sub, selection.method = "vst", nfeatures = 2000)

top10N <- head(VariableFeatures(dmd_sub), 10)
top10S <- head(VariableFeatures(cntl_sub), 10)

plot1N <- VariableFeaturePlot(dmd_sub)
plot2N <- LabelPoints(plot = plot1N, points = top10N, repel = TRUE, xnudge = 0, ynudge = 0)
plot1S <- VariableFeaturePlot(cntl_sub)
plot2S <- LabelPoints(plot = plot1S, points = top10S, repel = TRUE, xnudge = 0, ynudge = 0)

plot2N
plot2S

```

**Integration**

Integrating DMD and Control treatments. This corrects batch effect and allows cluster analysis of all data

```{r results='hide'}
objlist <- list(dmd_sub, cntl_sub)
features <- SelectIntegrationFeatures(object.list = objlist)
muscle.anchors <- FindIntegrationAnchors(object.list = objlist, anchor.features = features)
muscle.combined <- IntegrateData(anchorset = muscle.anchors)
```


Scale data, run a PCA, creat umap/cluster cells basd on expression profiles
```{r results='hide'}
DefaultAssay(muscle.combined) <- "integrated"
muscle.combined <- ScaleData(muscle.combined, verbose = FALSE)
muscle.combined <- RunPCA(muscle.combined, npcs = 7, verbose = FALSE)
muscle.combined <- RunUMAP(muscle.combined, reduction = "pca", dims = 1:7)
muscle.combined <- FindNeighbors(muscle.combined, reduction = "pca", dims = 1:7)
muscle.combined <- FindClusters(muscle.combined, resolution = 0.3)
```
#Check Dimensions to ensure we include as few as possible to reduce cluster number, including only significant dimensional factors
```{r}
#muscle.combined <- JackStraw(muscle.combined, num.replicate = 100)
#muscle.combined <- ScoreJackStraw(muscle.combined, dims = 1:8)

#JackStrawPlot(muscle.combined, dims = 1:8)
ElbowPlot(muscle.combined)
```

```{r results='hide'}
DefaultAssay(dmd_sub) <- "RNA"
dmd_sub <- ScaleData(dmd_sub, verbose = FALSE)
dmd_sub <- RunPCA(dmd_sub, npcs = 8, verbose = FALSE)
dmd_sub <- RunUMAP(dmd_sub, reduction = "pca", dims = 1:8)
dmd_sub <- FindNeighbors(dmd_sub, reduction = "pca", dims = 1:8)
dmd_sub <- FindClusters(dmd_sub, resolution = 0.3)
```
```{r results='hide'}
DefaultAssay(cntl_sub) <- "RNA"
cntl_sub <- ScaleData(cntl_sub, verbose = FALSE)
cntl_sub <- RunPCA(cntl_sub, npcs = 8, verbose = FALSE)
cntl_sub <- RunUMAP(cntl_sub, reduction = "pca", dims = 1:8)
cntl_sub <- FindNeighbors(cntl_sub, reduction = "pca", dims = 1:8)
cntl_sub <- FindClusters(cntl_sub, resolution = 0.3)
```

UMAP plotting. Pt.size is just the size of the dots. More cells = lower pt.size
```{r}
p0 = DimPlot(muscle.combined, reduction = "umap", split.by = "orig.ident", pt.size = 0.5)
p1 <- DimPlot(muscle.combined, reduction = "umap", group.by = "orig.ident", pt.size = 0.5)
p2 <- DimPlot(muscle.combined, reduction = "umap", label = TRUE, repel = TRUE, pt.size = 0.5)
p1 + p2
p2
p0
```
```{r}
p1 <- DimPlot(dmd_sub, reduction = "umap", group.by = "orig.ident", pt.size = 0.5)
p2 <- DimPlot(dmd_sub, reduction = "umap", label = TRUE, repel = TRUE, pt.size = 0.5)
p1 + p2
p3 <- DimPlot(cntl_sub, reduction = "umap", group.by = "orig.ident", pt.size = 0.5)
p4 <- DimPlot(cntl_sub, reduction = "umap", label = TRUE, repel = TRUE, pt.size = 0.5)
p3 + p4

```


Find conserved markers across per cluster the two data sets and saving them to a cluster variable. This ranks the most differentially expressed genes per integrated cluster ao that the list can be used to characterize cell clusters using GTEx or CellKb SC expression multi-gene sc query

```{r}
DefaultAssay(muscle.combined) <- "RNA"
cluster0.markers <- FindConservedMarkers(muscle.combined, ident.1 = 0, grouping.var = "orig.ident", verbose = FALSE)
cluster1.markers <- FindConservedMarkers(muscle.combined, ident.1 = 1, grouping.var = "orig.ident", verbose = FALSE)
cluster2.markers <- FindConservedMarkers(muscle.combined, ident.1 = 2, grouping.var = "orig.ident", verbose = FALSE)
cluster3.markers <- FindConservedMarkers(muscle.combined, ident.1 = 3, grouping.var = "orig.ident", verbose = FALSE)
cluster4.markers <- FindConservedMarkers(muscle.combined, ident.1 = 4, grouping.var = "orig.ident", verbose = FALSE)
cluster5.markers <- FindConservedMarkers(muscle.combined, ident.1 = 5, grouping.var = "orig.ident", verbose = FALSE)
cluster6.markers <- FindConservedMarkers(muscle.combined, ident.1 = 6, grouping.var = "orig.ident", verbose = FALSE)
cluster7.markers <- FindConservedMarkers(muscle.combined, ident.1 = 7, grouping.var = "orig.ident", verbose = FALSE)
cluster8.markers <- FindConservedMarkers(muscle.combined, ident.1 = 8, grouping.var = "orig.ident", verbose = FALSE)
cluster9.markers <- FindConservedMarkers(muscle.combined, ident.1 = 9, grouping.var = "orig.ident", verbose = FALSE)
cluster10.markers <- FindConservedMarkers(muscle.combined, ident.1 = 10, grouping.var = "orig.ident", verbose = FALSE)
```

Plot canonical cell-type gene marker expression to characterize cell clusters, confirm with GTEx Query of conserved cluster markers
```{r}

f1 = FeaturePlot(muscle.combined, split.by = 'orig.ident', features = c("TLR7"))
f1

#ggsave("/Users/ikegamik/Documents/kikegami/side_projects/tlr7_hunt/figs/homo/homo_tlr7_split_fp.png", plot = last_plot(), width = 7, height = 5, units = "in")
```

```{r}
FeaturePlot(muscle.combined, features = c("TLR7"))
#FeaturePlot(muscle.combined, features = c("NFKB1", 'CCL2', 'DMD', 'RELA'))
ggsave("/Users/ikegamik/Documents/kikegami/side_projects/tlr7_hunt/homo_int_tlr7.png", plot = last_plot(), width = 7, height = 5, units = "in")
```

```{r}
p10 = FeaturePlot(muscle.combined, features = c("PTPRC", 'TTN', 'ADIPOQ', 'DCN','PAX7', 'MYH11', 'PTPRB', 'MYH3'))
ggsave("/Users/ikegamik/Documents/kikegami/side_projects/tlr7_hunt/homo_celltyping.png", plot = last_plot(), width = 7, height = 5, units = "in")
```

```{R}
new_cluster_ids = c("Myofibers/blasts", 'Myofibers/blasts', 'Fibroblasts', 'Myofibers', 'Endothelial','Satellite', 'Immune' , 'Smooth muscle', 'Immune', 'Adipocytes', 'Myoblasts')
names(new_cluster_ids) <- levels(muscle.combined)
muscle.combined <- RenameIdents(muscle.combined, new_cluster_ids)
```

```{r}
DimPlot(muscle.combined, reduction = "umap", label = TRUE, pt.size = 0.5) + NoLegend()
ggsave("/Users/ikegamik/Documents/kikegami/side_projects/tlr7_hunt/homo_int_clusters.png", plot = last_plot(), width = 7, height = 5, units = "in")
```
```{r}
DimPlot(muscle.combined, reduction = "umap", split.by = "orig.ident", pt.size = 0.5)
ggsave("/Users/ikegamik/Documents/kikegami/side_projects/tlr7_hunt/homo_all_treat_clusters.png", plot = last_plot(), width = 7, height = 5, units = "in")
```

```{r}
#FeaturePlot(dmd_sub, features = c("TLR7"))
#FeaturePlot(dmd_sub, features = c('TTN', 'ADIPOQ', 'DCN', 'PAX7', 'MYH11', 'PTPRB', 'MYH3'))
#FeaturePlot(dmd_sub, features = c("NFKB1", 'CCL2', 'DMD', 'RELA'))
#FeaturePlot(dmd_sub, features = c('NFKB1'))
#FeaturePlot(dmd_sub, features = c("TLR2", 'TLR4', 'TLR7', 'TLR9'))
#VlnPlot(muscle.combined, group.by = "orig.ident", features = c("TLR7", 'NFKB1', 'DMD'))
#VlnPlot(dmd_sub, features = c("TLR7", 'NFKB1', 'DMD'))
VlnPlot(muscle.combined, group.by = "orig.ident", features = c("TLR7", 'NFKB1', 'IL1B', 'TNF', 'CCL2'))
ggsave("/Users/ikegamik/Documents/kikegami/side_projects/tlr7_hunt/homo_vln_plot.png", plot = last_plot(), width = 7, height = 5, units = "in")

```

```{r}
FeaturePlot(cntl_sub, features = c("TLR7", 'TTN', 'ADIPOQ', 'DCN', 'PAX7', 'MYH11', 'PTPRB', 'MYH3'))
FeaturePlot(cntl_sub, features = c("TLR2", 'TLR4', 'TLR7', 'TLR9'))
```
