---
title: "mouse_tlr7_hunt"
author: "Kobe Ikegami"
date: "2023-10-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(SeuratObject)
library(Seurat)
library(dplyr)
library(Seurat)
library(patchwork)
library(tidyverse)
```
#Load in data, create seurat objects for all treatments
```{r warning=FALSE}
#Import data and create objects: for each sample
c57bl.data <- Read10X(data.dir = "/Users/ikegamik/Documents/kikegami/side_projects/tlr7_hunt/c57bl")
mdx.data <- Read10X(data.dir = "/Users/ikegamik/Documents/kikegami/side_projects/tlr7_hunt/mdx")
e23aon.data <- Read10X(data.dir = "/Users/ikegamik/Documents/kikegami/side_projects/tlr7_hunt/e23aon")

c57bl <- CreateSeuratObject(counts = c57bl.data, project = "c57bl", min.cells = 3, min.features = 200)
mdx <- CreateSeuratObject(counts = mdx.data, project = "mdx", min.cells = 3, min.features = 200)
e23aon <- CreateSeuratObject(counts = e23aon.data, project = "e23aon", min.cells = 3, min.features = 200)
```

Searches for mitochondrial RNA based on the "Mt" tag (homo = 'MT', mus = 'mt')
```{r}
c57bl[["percent.mt"]] <- PercentageFeatureSet(c57bl, pattern = "^mt-")
mdx[["percent.mt"]] <- PercentageFeatureSet(mdx, pattern = "^mt-")
e23aon[["percent.mt"]] <- PercentageFeatureSet(e23aon, pattern = "^mt-")
```

#Plot raw count data for reads, features and % MT DNA for filtration of outlying cells
```{r}
#Pre-FILTER
plot1N <- FeatureScatter(c57bl, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2N <- FeatureScatter(mdx, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot3N <- FeatureScatter(e23aon, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot1N + plot2N + plot3N
plot1N <- FeatureScatter(c57bl, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot2N <- FeatureScatter(mdx, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot3N <- FeatureScatter(e23aon, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1N + plot2N + plot3N

VlnPlot(c57bl, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
VlnPlot(mdx, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
VlnPlot(e23aon, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```

```{r}
#APPLY FILTERS
c57bl <- subset(c57bl, subset = nFeature_RNA < 2000 & nFeature_RNA > 200 & nCount_RNA < 5000 & percent.mt < 8)
mdx <- subset(mdx, subset = nFeature_RNA < 2000 & nFeature_RNA > 200 & nCount_RNA < 5000 & percent.mt < 8)
e23aon <- subset(e23aon, subset = nFeature_RNA < 2000 & nFeature_RNA > 200 & nCount_RNA < 5000 & percent.mt < 8)

```

```{r}
#Filtered
plot1N <- FeatureScatter(c57bl, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2N <- FeatureScatter(mdx, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot3N <- FeatureScatter(e23aon, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot1N + plot2N + plot3N
plot1N <- FeatureScatter(c57bl, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot2N <- FeatureScatter(mdx, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot3N <- FeatureScatter(e23aon, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1N + plot2N

VlnPlot(c57bl, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
VlnPlot(mdx, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
VlnPlot(e23aon, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```
Normalize data, find DE genes within each treatment, plot them
```{r warning=FALSE}
c57bl <- NormalizeData(c57bl)
mdx <- NormalizeData(mdx)
e23aon <- NormalizeData(e23aon)

c57bl <- FindVariableFeatures(c57bl, selection.method = "vst", nfeatures = 2000)
mdx <- FindVariableFeatures(mdx, selection.method = "vst", nfeatures = 2000)
e23aon <- FindVariableFeatures(e23aon, selection.method = "vst", nfeatures = 2000)

top10N <- head(VariableFeatures(c57bl), 10)
top10S <- head(VariableFeatures(mdx), 10)
top10S <- head(VariableFeatures(e23aon), 10)

plot1N <- VariableFeaturePlot(c57bl)
plot2N <- LabelPoints(plot = plot1N, points = top10N, repel = TRUE, xnudge = 0, ynudge = 0)
plot1S <- VariableFeaturePlot(mdx)
plot2S <- LabelPoints(plot = plot1S, points = top10S, repel = TRUE, xnudge = 0, ynudge = 0)
plot1E <- VariableFeaturePlot(e23aon)
plot2E <- LabelPoints(plot = plot1S, points = top10S, repel = TRUE, xnudge = 0, ynudge = 0)

plot2N
plot2S
plot2E

```

**Integration**

Integrate all treatments into a single dataset, first starting with integration of control and treated disease mice

```{r results='hide'}
objlist <- list(e23aon, c57bl)
features <- SelectIntegrationFeatures(object.list = objlist)
muscle.anchors <- FindIntegrationAnchors(object.list = objlist, anchor.features = features)
muscle.combined <- IntegrateData(anchorset = muscle.anchors)
```
#integrate first int object with disease model mice
```{r results='hide'}
objlist <- list(muscle.combined, mdx)
features <- SelectIntegrationFeatures(object.list = objlist)
muscle.anchors <- FindIntegrationAnchors(object.list = objlist, anchor.features = features)
muscle.combined <- IntegrateData(anchorset = muscle.anchors)
```
#Elbow plot to determine appropriate number of dimensions
```{r}
#muscle.combined <- JackStraw(muscle.combined, num.replicate = 100)
#muscle.combined <- ScoreJackStraw(muscle.combined, dims = 1:5)

#JackStrawPlot(muscle.combined, dims = 1:8)
#ElbowPlot(muscle.combined)
```

```{r}
#ElbowPlot(muscle.combined)
#ElbowPlot(c57bl)
#ElbowPlot(mdx)
#ElbowPlot(e23aon)
```

Runs PCA, prepares data for UMAP and uses nearest neighbor algorithm to help with clustering.
```{r results='hide'}
DefaultAssay(muscle.combined) <- "integrated"
muscle.combined <- ScaleData(muscle.combined, verbose = FALSE)
muscle.combined <- RunPCA(muscle.combined, npcs = 5, verbose = FALSE)
muscle.combined <- RunUMAP(muscle.combined, reduction = "pca", dims = 1:5)
muscle.combined <- FindNeighbors(muscle.combined, reduction = "pca", dims = 1:5)
muscle.combined <- FindClusters(muscle.combined, resolution = 0.3)

```

```{r results='hide'}
DefaultAssay(c57bl) <- "RNA"
c57bl <- ScaleData(c57bl, verbose = FALSE)
c57bl <- RunPCA(c57bl, npcs = 4, verbose = FALSE)
c57bl <- RunUMAP(c57bl, reduction = "pca", dims = 1:4)
c57bl <- FindNeighbors(c57bl, reduction = "pca", dims = 1:4)
c57bl <- FindClusters(c57bl, resolution = 0.3)

```

```{r results='hide'}
DefaultAssay(mdx) <- "RNA"
mdx <- ScaleData(mdx, verbose = FALSE)
mdx <- RunPCA(mdx, npcs = 4, verbose = FALSE)
mdx <- RunUMAP(mdx, reduction = "pca", dims = 1:4)
mdx <- FindNeighbors(mdx, reduction = "pca", dims = 1:4)
mdx <- FindClusters(mdx, resolution = 0.3)

```

```{r results='hide'}
DefaultAssay(e23aon) <- "RNA"
e23aon <- ScaleData(e23aon, verbose = FALSE)
e23aon <- RunPCA(e23aon, npcs = 4, verbose = FALSE)
e23aon <- RunUMAP(e23aon, reduction = "pca", dims = 1:4)
e23aon <- FindNeighbors(e23aon, reduction = "pca", dims = 1:4)
e23aon <- FindClusters(e23aon, resolution = 0.3)

```
UMAP plotting. Pt.size is just the size of the dots. More cells = lower pt.size
```{r}
p1 <- DimPlot(muscle.combined, reduction = "umap", group.by = "orig.ident", pt.size = 0.5)
p2 <- DimPlot(muscle.combined, reduction = "umap", label = TRUE, repel = TRUE, pt.size = 0.5)
p1 + p2
p2
```

```{r}
p7 <- DimPlot(c57bl, reduction = "umap", group.by = "orig.ident", pt.size = 0.5)
p8 <- DimPlot(c57bl, reduction = "umap", label = TRUE, repel = TRUE, pt.size = 0.5)
p7 + p8
p3 <- DimPlot(mdx, reduction = "umap", group.by = "orig.ident", pt.size = 0.5)
p4 <- DimPlot(mdx, reduction = "umap", label = TRUE, repel = TRUE, pt.size = 0.5)
p3 + p4
p5 <- DimPlot(e23aon, reduction = "umap", group.by = "orig.ident", pt.size = 0.5)
p6 <- DimPlot(e23aon, reduction = "umap", label = TRUE, repel = TRUE, pt.size = 0.5)
p5 + p6

```

```{r}
DimPlot(muscle.combined, reduction = "umap", split.by = "orig.ident", pt.size = 0.5)
```
Find most DE genes per cluster in all treatments of integrated dataset. Use these to verify canonical gene marker expression in featureplots to ascertain cluster cell-type

```{r}
DefaultAssay(muscle.combined) <- "RNA"
cluster0.markers <- FindConservedMarkers(muscle.combined, ident.1 = 0, grouping.var = "orig.ident", verbose = FALSE)
cluster1.markers <- FindConservedMarkers(muscle.combined, ident.1 = 1, grouping.var = "orig.ident", verbose = FALSE)
cluster2.markers <- FindConservedMarkers(muscle.combined, ident.1 = 2, grouping.var = "orig.ident", verbose = FALSE)
cluster3.markers <- FindConservedMarkers(muscle.combined, ident.1 = 3, grouping.var = "orig.ident", verbose = FALSE)
cluster4.markers <- FindConservedMarkers(muscle.combined, ident.1 = 4, grouping.var = "orig.ident", verbose = FALSE)
cluster5.markers <- FindConservedMarkers(muscle.combined, ident.1 = 5, grouping.var = "orig.ident", verbose = FALSE)
cluster6.markers <- FindConservedMarkers(muscle.combined, ident.1 = 6, grouping.var = "orig.ident", verbose = FALSE)
cluster7.markers <- FindConservedMarkers(muscle.combined, ident.1 = 7, grouping.var = "orig.ident", verbose = FALSE)
cluster8.markers <- FindConservedMarkers(muscle.combined, ident.1 = 8, grouping.var = "orig.ident", verbose = FALSE)
cluster9.markers <- FindConservedMarkers(muscle.combined, ident.1 = 9, grouping.var = "orig.ident", verbose = FALSE)
cluster10.markers <- FindConservedMarkers(muscle.combined, ident.1 = 10, grouping.var = "orig.ident", verbose = FALSE)

```

## Characterizing Clusters with canonical gene-celltype markers

```{r}
f1 = FeaturePlot(muscle.combined, split.by="orig.ident", features = c("Tlr7"))
f1
ggsave("/Users/ikegamik/Documents/kikegami/side_projects/tlr7_hunt/figs/mus/mus_tlr7_split_fp.png", plot = last_plot(), width = 7, height = 5, units = "in")
```

```{r}
p2
f1 = FeaturePlot(muscle.combined, features = c("Tlr7"))
#ggsave("/Users/ikegamik/Documents/kikegami/side_projects/tlr7_hunt/mus_int_tlr7.png", plot = last_plot(), width = 7, height = 5, units = "in")
#FeaturePlot(muscle.combined, features = c("Ptprc", 'Ttn', 'Adipoq', 'Dcn'))
#FeaturePlot(muscle.combined, features = c('Pax7', 'Myh11', 'Ptprb', 'Myh3'))
#FeaturePlot(muscle.combined, features = c("Nfkb1", 'Ccl2', 'Dmd', 'Rela'))
#FeaturePlot(muscle.combined, features = c('Lbp', 'Hcam', 'Siglec3', 'Ia2b'))
FeaturePlot(muscle.combined, features = c("Ptprc", 'Ttn', 'Adipoq', 'Dcn','Pax7', 'Myh11', 'Ptprb', 'Myh3'))
ggsave("/Users/ikegamik/Documents/kikegami/side_projects/tlr7_hunt/mus_celltyping.png", plot = last_plot(), width = 7, height = 5, units = "in")
```
```{r}
new_cluster_ids = c("Myofibers", 'Fibroblasts', 'Myofibers',  'Myoblasts', 'Immune', 'Endothelial (vas)', 'Smooth muscle', 'Satellite', 'Adipocytes',  'Endothelial(vas)','Endothelial (lym)')
names(new_cluster_ids) <- levels(muscle.combined)
muscle.combined <- RenameIdents(muscle.combined, new_cluster_ids)
```

```{r}
p4
FeaturePlot(mdx, features = c("Ptprc", 'Ttn', 'Adipoq', 'Dcn','Pax7', 'Myh11', 'Ptprb', 'Myh3'))
FeaturePlot(mdx, features = c("Tlr7"))
```
```{r}
mdx_cluster_ids = c("Myofibers", 'Myofibers', 'Fibroblasts',  'Immune', 'Fibroblasts', 'Immune', 'Myofibers', 'Myoblasts/Sattelite', 'Endothelial','Smooth Muscle')
names(mdx_cluster_ids) <- levels(mdx)
mdx <- RenameIdents(mdx, mdx_cluster_ids)
```

```{r}
# find markers for every cluster compared to all remaining cells, report only the positive
# ones
c57bl.markers <- FindAllMarkers(c57bl, only.pos = TRUE)
c57bl.markers %>%
    group_by(cluster) %>%
    dplyr::filter(avg_log2FC > 1)
```

```{r}
p8
FeaturePlot(c57bl, features = c("Ptprc", 'Ttn', 'Adipoq', 'Dcn','Pax7', 'Myh11', 'Ptprb', 'Myh3'))
FeaturePlot(c57bl, features = c("Tlr7"))
```
```{r}
c57bl_cluster_ids = c("Myofibers", 'Myofibers', 'Fibroblasts', 'Endothelial', 'Myofibers','Smooth Muscle', 'Myoblasts/Sattelites', 'Immune','Myofibers','Adipocytes')
names(c57bl_cluster_ids) <- levels(c57bl)
c57bl <- RenameIdents(c57bl, c57bl_cluster_ids)
```

```{r}
DimPlot(muscle.combined, reduction = "umap", label = TRUE, pt.size = 0.5) + NoLegend()
#ggsave("/Users/ikegamik/Documents/kikegami/side_projects/tlr7_hunt/mus_int_clusters.png", plot = last_plot(), width = 7, height = 5, units = "in")
```

```{r}
DimPlot(muscle.combined, reduction = "umap", split.by = "orig.ident", pt.size = 0.5)
ggsave("/Users/ikegamik/Documents/kikegami/side_projects/tlr7_hunt/mus_all_treat_clusters.png", plot = last_plot(), width = 7, height = 5, units = "in")
```

```{r}
DimPlot(c57bl, reduction = "umap", label = TRUE, pt.size = 0.5) + NoLegend()

```

```{r}
#FeaturePlot(c57bl, features = c("Tlr7"))
FeaturePlot(c57bl, features = c("Tlr7"))
FeaturePlot(c57bl, features = c("Ptprc", 'Ttn', 'Adipoq', 'Dcn'))
FeaturePlot(c57bl, features = c('Pax7', 'Myh11', 'Ptprb', 'Myh3'))
FeaturePlot(c57bl, features = c("Nfkb1", 'Ccl2', 'Dmd', 'Rela'))
FeaturePlot(c57bl, features = c("Tlr2", 'Tlr4', 'Tlr7', 'Tlr9'))
VlnPlot(muscle.combined, group.by = "orig.ident", features = c("Tlr7", 'Nfkb1', 'Dmd'))
VlnPlot(c57bl, features = c("Tlr7", 'Nfkb1', 'Dmd'))
```

```{r}
mdx2 = mdx
c57bl2 = c57bl
#mdx2$cell_type = c("Myofibers", 'Myofibers', 'Fibroblasts',  'Immune', 'Fibroblasts', 'Immune', 'Myofibers', 'Myoblasts/Sattelite', 'Endothelial','Smooth Muscle')
#c57bl2$cell_type = c("Myofibers", 'Myofibers', 'Fibroblasts', 'Endothelial', 'Myofibers','Smooth Muscle', 'Myoblasts/Sattelites', 'Immune','Myofibers','Adipocytes')
mdx2_levels = c("Myofibers", 'Fibroblasts',  'Immune',  'Myoblasts/Sattelite', 'Endothelial','Smooth Muscle')
c57bl2_levels = c("Myofibers", 'Fibroblasts',  'Immune', 'Myoblasts/Sattelites','Endothelial', 'Smooth Muscle', 'Adipocytes')

Idents(mdx2) = factor(Idents(mdx2), levels= mdx2_levels)
Idents(c57bl2) = factor(Idents(c57bl2), levels= c57bl2_levels)

#mdx2$mdx_levels = factor(mdx2$seurat_clusters, levels= mdx2_levels)
#c57bl2$c57bl2_levels = factor(c57bl2$seurat_clusters, levels= c57bl2_levels)
```

```{r}
#VlnPlot(mdx, features = c("Tlr7", 'Nfkb1', 'Dmd'))
#VlnPlot(c57bl, features = c("Tlr7", 'Nfkb1', 'Dmd'))
v10 = VlnPlot(mdx2, features = c('Nfkb1'))
v11 = VlnPlot(c57bl2, features = c('Nfkb1'))
v10
v11
#muscle.combined$celltype.stim <- paste(Idents(muscle.combined), muscle.combined$sample, sep = "_") #assign the same on a new colunm
#muscle.combined$celltype <- Idents(muscle.combined) # assign the identify to the cell type colunm
#VlnPlot(muscle.combined, group.by = "orig.ident", features = c("Tlr7", 'Nfkb1', 'Dmd'))
```

```{r}
VlnPlot(c57bl2, features = c("Tlr7"))
ggsave("/Users/ikegamik/Documents/kikegami/side_projects/tlr7_hunt/mus_cntl_tlr7_vln-plot.png", plot = last_plot(), width = 7, height = 5, units = "in")
VlnPlot(c57bl2, features = c('Nfkb1'))
ggsave("/Users/ikegamik/Documents/kikegami/side_projects/tlr7_hunt/mus_cntl_nfkb1_vln-plot.png", plot = last_plot(), width = 7, height = 5, units = "in")
VlnPlot(c57bl2, features = c('Tnf'))
ggsave("/Users/ikegamik/Documents/kikegami/side_projects/tlr7_hunt/mus_cntl_tnf_vln-plot.png", plot = last_plot(), width = 7, height = 5, units = "in")
VlnPlot(c57bl2, features = c('Ccl2'))
ggsave("/Users/ikegamik/Documents/kikegami/side_projects/tlr7_hunt/mus_cntl_ccl2_vln-plot.png", plot = last_plot(), width = 7, height = 5, units = "in")
```
```{r}
VlnPlot(mdx2, features = c("Tlr7"))
ggsave("/Users/ikegamik/Documents/kikegami/side_projects/tlr7_hunt/mus_mdx_tlr7_vln-plot.png", plot = last_plot(), width = 7, height = 5, units = "in")
VlnPlot(mdx2, features = c('Nfkb1'))
ggsave("/Users/ikegamik/Documents/kikegami/side_projects/tlr7_hunt/mus_mdx_nfkb1_vln-plot.png", plot = last_plot(), width = 7, height = 5, units = "in")
VlnPlot(mdx2, features = c('Tnf'))
ggsave("/Users/ikegamik/Documents/kikegami/side_projects/tlr7_hunt/mus_mdx_tnf_vln-plot.png", plot = last_plot(), width = 7, height = 5, units = "in")
VlnPlot(mdx2, features = c('Ccl2'))
ggsave("/Users/ikegamik/Documents/kikegami/side_projects/tlr7_hunt/mus_mdx_ccl2_vln-plot.png", plot = last_plot(), width = 7, height = 5, units = "in")
```

```{r}
FeaturePlot(mdx, features = c("Tlr7"))
FeaturePlot(mdx, features = c("Ptprc", 'Ttn', 'Adipoq', 'Dcn'))
FeaturePlot(mdx, features = c('Pax7', 'Myh11', 'Ptprb', 'Myh3'))
FeaturePlot(mdx, features = c("Nfkb1", 'Ccl2', 'Dmd', 'Rela'))
FeaturePlot(mdx, features = c("Tlr2", 'Tlr4', 'Tlr7', 'Tlr9'))
```
```{r}
FeaturePlot(e23aon, features = c("Tlr7"))
FeaturePlot(e23aon, features = c("Ptprc", 'Ttn', 'Adipoq', 'Dcn'))
FeaturePlot(e23aon, features = c('Pax7', 'Myh11', 'Ptprb', 'Myh3'))
FeaturePlot(e23aon, features = c("Nfkb1", 'Ccl2', 'Dmd', 'Rela'))
FeaturePlot(e23aon, features = c("Tlr2", 'Tlr4', 'Tlr7', 'Tlr9'))
```

```{r}
cluster2.markers <- FindMarkers(muscle.combined, ident.1 = 2, min.pct = 0.25)
head(cluster2.markers, n = 5)
cluster2.markers <- FindMarkers(muscle.combined, ident.1 = 7, min.pct = 0.25)
head(cluster2.markers, n = 5)
```

```{r}
(length(WhichCells(mdx2, expression = Tlr7 > 0, ident = 'Immune'))/length(which(Idents(mdx2) == 'Immune'))) *100
(length(WhichCells(mdx2, expression = Tlr7 > 0, ident = 'Myoblasts/Sattelite'))/length(which(Idents(mdx2) == 'Myoblasts/Sattelite')))*100
(length(WhichCells(mdx2, expression = Tlr7 > 0, ident = 'Myofibers'))/length(which(Idents(mdx2) == 'Myofibers')))*100
(length(WhichCells(mdx2, expression = Tlr7 > 0, ident = 'Fibroblasts'))/length(which(Idents(mdx2) == 'Fibroblasts')))*100
(length(WhichCells(c57bl2, expression = Tlr7 > 0, ident = 'Immune'))/length(which(Idents(c57bl2) == 'Immune')))*100
(length(WhichCells(c57bl2, expression = Tlr7 > 0, ident = 'Fibroblasts'))/length(which(Idents(c57bl2) == 'Fibroblasts')))*100
```